### 2020-02-19 thesis/content/chapter-related-work.tex
continued wl section

To classify or score a graph, it first needs to be characterized by a set of relevant properties.
Two commonly used graph characterization approaches are the so called \acl{wl} coloring and the notion of the graph spectrum.
Both will now be briefly introduced since they form the foundation of most current \ac{gcr} methods.

@@

The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
A mapping $\chi_{G, k}: \mathcal{V}^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G = (\mathcal{V}, \mathcal{E})$.
We say  $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.

The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^0 \preceq \chi_{G, k}^1 \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^i \equiv \chi_{G, k}^{i+1}$ is satisfied;
here the equivalence operator $\equiv$ denotes that $\chi_{G, k}^i$ and $\chi_{G, k}^{i+1}$ must be identical up to color/label substitutions.
The final \ac{wl} coloring $\bar{\chi}_{G, k}$ that is obtained after convergence can be used to quickly check whether two graphs $G, H$ are not isomorphic ($G \not\simeq H$).
\begin{defn}\label{defn:related:wl-distinguishable}
	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \bar{\chi}_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \bar{\chi}_{H, k}(b) = c \right\}\right|$. % chktex 21
\end{defn}
It is easy to show that $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
The opposite does however not necessarily hold;
two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \centernot\implies G \simeq H$.
To understand why this is the case and which types of graphs are $k$-\acs{wl} distinguishable, we will now look at the details of \ac{wl} color refinement step.
First the color refinement algorithm for the most simple case of $k = 1$ is described.
Then the definitions and intuitions from the 1-dimensional case are extended to its higher-dimensional generalization.
Lastly we will discuss the discriminative power of the \acs{wl} algorithm and its relation to the \acs{wl}-dimension $k$.

\subsubsection{The 1-dimensional \acs{wl} algorithm}
In the 1-dimensional \ac{wl} algorithm, a color is assigned to each vertex of a graph.
If the vertices $\mathcal{V}_G$ of the input graph $G$ are already labeled via some labeling $l_G: \mathcal{V}_G \to L$, those labels can be used as the initial graph coloring $\chi_{G,1}^0 = l_G$.
For unlabeled graphs a constant coloring is used instead, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^0(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$. % chktex 25
In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
\begin{defn}\label{defn:related:wl1-refine}
	$\chi_{G,1}^{i+1}(v) \coloneqq h\left(\chi_{G,1}^{i}(v), \ldblbrace \chi_{G,1}^{i}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
	with $\Gamma_G(v)$ denoting the set of adjecent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
	In practice $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
\end{defn}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\linewidth]{gfx/related-work/wl1-refine.pdf}
	\caption[Example 1-WL color refinement steps.]{
		Example 1-WL color refinement steps.
		After two iterations the coloring stabilizes.
		Each vertex is labeled with its current color and has the aggregated neighbors, as in \cref{defn:related:wl1-refine}, written next to it.
	}\label{fig:related:wl1-refine}
\end{figure}

\subsubsection{The $k$-dimensional \acs{wl} algorithm}
As we just saw, the 1-\acs{wl} algorithm iteratively refines colorings of single vertices.
While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the graph isomorphism problem as illustrated in \cref{fig:related:wl1-problem}.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl1-problem.pdf}
	\caption[Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.]{
		Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.
	}\label{fig:related:wl1-problem}
\end{figure}
By extending \ac{wl} to higher dimensions such 1-\acs{wl} indistinguishable cases can however be handled.
Analogous to the 1-dimensional \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
\begin{defn}\label{defn:related:wlk-refine}
	$\begin{aligned}[t]
		\chi_{G,k}^{i+1}(v) \coloneqq &\ h\left(\chi_{G,k}^{i}(v), \ldblbrace (\chi_{G,k}^{i}(v[u/1]), \dots, \chi_{G,k}^{i}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
		\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}^k, v[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
	\end{aligned}$
\end{defn}
In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
In $k$-\acs{wl} the color of a $k$-tuple $v \in \mathcal{V}^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $v$.
Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
This more abstract notion of neighborhood is illustrated in \cref{fig:related:wl-neighbors}.
\begin{figure}[ht]
	\centering
	\begin{subfigure}{0.33\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl1-neighbors.pdf}
		\subcaption{1-WL $v_1$ neighbors}\label{fig:related:wl-neighbors:1}
	\end{subfigure}%
	\begin{subfigure}{0.33\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl2-neighbors.pdf}
		\subcaption{2-WL $(v_1, v_2)$ neighbors}\label{fig:related:wl-neighbors:2}
	\end{subfigure}%
	\begin{subfigure}{0.33\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl3-neighbors.pdf}
		\subcaption{3-WL $(v_1, v_2, v_3)$ neighbors}\label{fig:related:wl-neighbors:3}
	\end{subfigure}
	\caption[\ac{wl} neighborhoods for different values of $k$.]{
		Tuple neighborhoods for different values of $k$.
		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple $v \in \mathcal{V}_G^k$ whose neighbors are shown;
		for simplicity neighbors with $u \in v$ are left out (see~\cref{defn:related:wlk-refine}).
		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_green}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
	}\label{fig:related:wl-neighbors}
\end{figure}
For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible paths of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\bar{\chi}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.

Let us now look at how the tuple colors are initialized.
For this we use subgraph \textit{isomorphism types} which determine the initial color $\chi_{G,k}^0(v)$ of each $k$-tuple $v$.
For $k = 1$ the isomorphism type of a vertex $v$ directly corresponds to its label $l_G(v)$.
For $k > 1$ the isomorphism types more generally correspond to the classes of isomorphic ordered subgraphs of $G$ induced by the $k$-tuples $\mathcal{V}_G^k$.
Formally this means that
\begin{align}
	\chi_{G,k}^0(v_1, \dots, v_k) = \chi_{G,k}^0(w_1, \dots, w_k) \label{eq:related:wlk-init}
	\iff &\ \forall i: l_G(v_i) = l_G(w_i) \\
	\land &\ \forall i, j: v_i = v_j \leftrightarrow w_i = w_j \nonumber \\
	\land &\ \forall i, j: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_G \text{.} \nonumber
\end{align}
If no explicit vertex labeling $l_G$ is given, a single constant label for all vertices is assumed.
Also note that there is a fundamental difference in how the adjacency information encoded in $\mathcal{E}_G$ is used in 1-\acs{wl} vs.\ $k$-\acs{wl}:
In 1-\acs{wl} a vertex coloring by itself cannot encode adjacency which is why this information is explicitly incorporated in each refinement step via $\Gamma_G$ (see~\cref{defn:related:wl1-refine}).
In $k$-\acs{wl} on the other hand each pair of vertices $(v, u) \in \mathcal{V}_G^2$ appears in at least one $k$-tuple (assuming $k \geq 2$) and therefore has at least one color which can implicitly encode the adjacency information.
Edges and non-edges are colored differently in the initial coloring as defined in \cref{eq:related:wlk-init};
thus no explicit adjacency information is needed in the $k$-\acs{wl} color refinement step in \cref{defn:related:wlk-refine}.

\subsubsection{Discriminative Power of \acs{wl}}
Now we will look at the types of graphs that can be distinguished by \acs{wl} in relation to the \acs{wl}-dimension $k$.
\begin{lem}\label{lem:related:wl-lower}
	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$ ($(k+1)$-\acs{wl} is at least as powerful as $k$-\acs{wl}).
\end{lem}
\begin{hproof}
	All $k$-tuples can be mapped to $(k+1)$-tuples via $\varphi(v_1, \dots, v_k) \coloneqq (v_1, \dots,\allowbreak v_k, v_k)$.
	For each neighbor $(v[u/1], \dots,\allowbreak v[u/k])$ of $v$, there is a corresponding neighbor $(\varphi(v[u/1]), \dots \varphi(v[u/k]), \varphi(v[u/k]))$ of $\varphi(v)$.
	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall v, w \in \mathcal{V}_G^k: \chi_{G, k}^i(v) \neq \chi_{G, k}^i(w) \rightarrow \chi_{G, k+1}^i(\varphi(v)) \neq \chi_{G, k+1}^i(\varphi(w))$.
	The lemma then follows by \cref{defn:related:wl-distinguishable}.
\end{hproof}
\begin{fact}[see~\citet{Immerman1990}]\label{fact:related:wl-upper}
	For all $k$ there are non-isomorphic graphs $G \not\simeq H$ of size $\mathcal{O}(k)$ with $G \mathrel{\simeq_k} H$.
\end{fact}
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl2-problem-solution.pdf}
	\caption{
		Two non-isomorphic graphs with $G \mathrel{\simeq_1} H$ and $G \mathrel{\not\simeq_2} H$.
	}\label{fig:related:wl2-problem-solution}
\end{figure}
\begin{cor}\label{cor:related:wl-monotonous}
	The discriminative power of $k$-\acs{wl} grows monotonously with $k$ and never converges, i.e.\ for all $k \in \mathbb{N}$ there is an $l \in \mathbb{N}$ s.t.\ the set of $k$-\ac{wl} distinguishable graphs is a true subset of the $(k+l)$-\ac{wl} distinguishable graphs.
\end{cor}
\begin{hproof}
	Note that $k$-\acs{wl} trivially solves the graph isomorphism problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
	Using \cref{lem:related:wl-lower} and \cref{fact:related:wl-upper} the corollary directly follows.
\end{hproof}
\Cref{fig:related:wl2-problem-solution} illustrates \cref{cor:related:wl-monotonous} by showing how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.
Since the time complexity of $k$-\acs{wl} grows exponentially with $k$~\cite[corr.~1.9.7]{Immerman1990}, a trade-off between the runtime and the discriminative power has to be made.

@@

\paragraph{Graph Embeddings}
Skip-gram methods can not only be used for vertex embeddings but also to embed entire graphs.
One way to do this is via the \texttt{graph2vec}~\cite{Narayanan2017} method.
It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
\texttt{graph2vec} gets a set of graphs $\mathbb{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $f(G_1), \dots, f(G_N)$.
While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathbb{G}$ as context roots.
The context $C_k(G_i)$ of a graph $G_i = (\mathcal{V}_i, \mathcal{E}_i)$ is defined as
\begin{equation}
	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_i}}\, {\left\{ {\chi_{G_i, 1}^l(v_j) \right\}}_{l = 0}^k % chktex 21
	\text{ with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^l$ as in \cref{defn:related:wl1-refine}.}
\end{equation}
Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2k$.
Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
Using the previous definitions, the context root embedding function has the signature $f: \mathbb{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathbb{G}} C_k(G_i) \to \mathbb{R}^d$.
To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:gcr:embed:word2vec} is reused.
Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.

>>>>

\begin{defn}
	A graph $G \coloneqq (\mathcal{V}_G, \mathcal{E}_G)$ consists of a set of vertices $v_i \in \mathcal{V}_G$ and a set of edges $e_{ij} = (v_i, v_j) \in \mathcal{E}_G \subseteq \mathcal{V}_G^2$.
	Optionally discrete vertex labels $l_i \in L_{\mathcal{V}}$ or edge labels $l_{ij} \in L_{\mathcal{E}}$ may be associated with all vertices $v_i \in \mathcal{V}_G$ and edges $e_{ij}\in \mathcal{E}_G$ respectively.
	Also continuous feature vectors $x_{i} \in \mathcal{X}_{\mathcal{V}}$, $x_{ij} \in \mathcal{X}_{\mathcal{E}}$ may be given.
	If $\mathcal{X}_{\mathcal{E}} = \mathbb{R}$, $x_{ij}$ can for example be interpreted as an edge weight of $e_{ij}$.
\end{defn}
In this thesis only undirected graphs $G$ will be considered, i.e.\ ${e_{ij} \in \mathcal{E}_G} \leftrightarrow\allowbreak {e_{ji} \in \mathcal{E}_G} \allowbreak\land l_{ij} = l_{ji} \allowbreak\land x_{ij} = x_{ji}$.
To classify or score such a graph, it first needs to be characterized by a set of relevant properties.
The most strict characterization of a graph is its so-called \textit{isomorphism type}.
It uniquely identifies a graph but lacks any notion of similarity between non-identical graphs.
We will begin with a brief definition of this strict isomorphism-based graph characterization.
Then two less strict characterization approaches are described; the so called \acl{wl} coloring and the notion of graph spectra.
They are the theoretical foundation of many current \ac{gcr} methods.

\subsection{The Graph Isomorphism Problem}%
\label{sec:related:character:iso}

@@

The \acf{wl} algorithm~\cite{Weisfeiler1968}\cite{Cai1992} characterizes a graph $G$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}_G^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
A mapping $\chi_{G, k}: \mathcal{V}_G^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G$.
\begin{defn}
	A coloring $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}_G^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.
\end{defn}
\begin{defn}
	Two colorings $\chi$ and $\chi'$ are \textit{equivalent} ($\chi \equiv \chi'$) iff.\ $\forall a, b \in \mathcal{V}_G^k: {\chi(a) = \chi(b)} \leftrightarrow {\chi'(a) = \chi'(b)}$, i.e.\ $\chi$ is identical to $\chi'$ up to color substitutions.
\end{defn}
The $k$-dimensional \ac{wl} algorithm ($k$-\acs{wl}) works by iteratively refining $k$-colorings $\chi_{G, k}^{(0)} \preceq \chi_{G, k}^{(1)} \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^{(i)} \equiv \chi_{G, k}^{(i+1)}$ is satisfied.
We denote the final, maximally refined, $k$-\acs{wl} coloring with $\chi^{*}_{G, k}$.
\begin{defn}\label{defn:related:wl-distinguishable}
	Two graphs $G$ and $H$ are are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) iff.\ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \chi^{*}_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \chi^{*}_{H, k}(b) = c \right\}\right|$. % chktex 21
\end{defn}
It is easy to show that $G \mathrel{{\not\simeq}_k} H \implies G \not\simeq H$.
The opposite does however not necessarily hold;
two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \centernot\implies G \simeq H$.
To understand why this is the case and which types of graphs are $k$-\acs{wl} distinguishable, we will now look at the details of \ac{wl} color refinement step.
First the color refinement algorithm for the most simple case of $k = 1$ is described.
Then the definitions and intuitions from the 1-dimensional case are extended to its higher-dimensional generalization.
Lastly we will discuss the discriminative power of the \acs{wl} algorithm and its relation to the \acs{wl}-dimension $k$.

\subsubsection{The 1-dimensional \acs{wl} algorithm}
In the 1-dimensional \ac{wl} algorithm, a color is assigned to each vertex of a graph.
If the vertices $\mathcal{V}_G$ of the input graph $G$ are already labeled via some labeling $l_G: \mathcal{V}_G \to L$, those labels can be used as the initial graph coloring $\chi_{G,1}^{(0)} = l_G$.
For unlabeled graphs a constant coloring is used instead, e.g.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^{(0)}(v) = \textcolor{t_blue}{\texttt{A}}$ for some initial color $\textcolor{t_blue}{\texttt{A}} \in \mathcal{C}$. % chktex 25
In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
\begin{defn}\label{defn:related:wl1-refine}
	$\chi_{G,1}^{(i+1)}(v) \coloneqq h\left(\chi_{G,1}^{(i)}(v), \ldblbrace \chi_{G,1}^{(i)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)$,
	with $\Gamma_G(v)$ denoting the set of adjecent vertices of $v \in \mathcal{V}_G$ and $h: \mathcal{C}^* \to \mathcal{C}$ denoting a perfect hash function that assigns a unique color to each finite combination of colors.
	In practice $h$ is usually defined lazily by using $\mathcal{C} = \mathbb{N}$ and non-deterministically enumerating color combinations in the order in which they are hashed s.t.\ a new color is introduced every time a previously unseen color combination is queried.
\end{defn}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\linewidth]{gfx/related-work/wl1-refine.pdf}
	\caption[Example 1-WL color refinement steps.]{
		Example 1-WL color refinement steps.
		After two iterations the coloring stabilizes.
		Each vertex is labeled with its current color and has the aggregated neighbors, as in \cref{defn:related:wl1-refine}, written next to it.
	}\label{fig:related:wl1-refine}
\end{figure}

\subsubsection{The $k$-dimensional \acs{wl} algorithm}
As we just saw, the 1-\acs{wl} algorithm iteratively refines colorings of single vertices.
While the obtained colorings differ for most non-isomorphic graphs $G \not\simeq H$, 1-\acs{wl} does not generally solve the graph isomorphism problem as illustrated in \cref{fig:related:wl1-problem}.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl1-problem.pdf}
	\caption[Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.]{
		Two simple non-isomorphic graphs that are indistinguishable by 1-\acs{wl}.
	}\label{fig:related:wl1-problem}
\end{figure}
By extending \ac{wl} to higher dimensions such 1-\acs{wl} indistinguishable cases can however be handled.
Analogous to the 1-dimensional \cref{defn:related:wl1-refine}, the $k$-dimensional color refinement step is defined by
\begin{defn}\label{defn:related:wlk-refine}
	$\begin{aligned}[t]
		\chi_{G,k}^{(i+1)}(v) \coloneqq &\ h\left(\chi_{G,k}^{(i)}(v), \ldblbrace (\chi_{G,k}^{(i)}(v[u/1]), \dots, \chi_{G,k}^{(i)}(v[u/k]))\, |\, u \in \mathcal{V}_G \rdblbrace\right)\\
		\text{with } v = &\ (v_1, \dots, v_k) \in \mathcal{V}_G^k, v[u/j] \coloneqq (v_1, \dots, v_{j-1}, u, v_{j+1}, \dots, v_k) \text{.}
	\end{aligned}$
\end{defn}
In 1-\acs{wl} a vertex color is refined by combining the colors of neighboring vertices.
In $k$-\acs{wl} the color of a $k$-tuple $v \in \mathcal{V}_G^k$ is refined by combining the colors of its neighborhood which is defined as the set of all $k$-tuples in which at most one vertex differs from $v$.
Note that each vertex $k$-tuple has one neighbor for each $u \in \mathcal{V}_G$, each of which is a $k$-tuple of vertex $k$-tuples.
This more abstract notion of neighborhood is illustrated in \cref{fig:related:wl-neighbors}.
\begin{figure}[ht]
	\centering
	\begin{subfigure}{0.33\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl1-neighbors.pdf}
		\subcaption{1-WL $v_1$ neighbors}\label{fig:related:wl-neighbors:1}
	\end{subfigure}%
	\begin{subfigure}{0.33\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl2-neighbors.pdf}
		\subcaption{2-WL $(v_1, v_2)$ neighbors}\label{fig:related:wl-neighbors:2}
	\end{subfigure}%
	\begin{subfigure}{0.33\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{gfx/related-work/wl3-neighbors.pdf}
		\subcaption{3-WL $(v_1, v_2, v_3)$ neighbors}\label{fig:related:wl-neighbors:3}
	\end{subfigure}
	\caption[\ac{wl} neighborhoods for different values of $k$.]{
		Tuple neighborhoods for different values of $k$.
		The vertices highlighted in \textcolor{t_blue}{blue} form the root tuple $v \in \mathcal{V}_G^k$ whose neighbors are shown;
		for simplicity neighbors with $u \in v$ are left out (see~\cref{defn:related:wlk-refine}).
		Each neighbor is highlighted with a different color, except for 3-WL where the \textcolor{t_red}{red}, \textcolor{t_green}{green} and \textcolor{t_darkyellow}{yellow} triples actually form the single neighbor for $u = v_4$.
	}\label{fig:related:wl-neighbors}
\end{figure}
For $k = 2$ this means that each potential edge $(v, w) \in \mathcal{V}_G^2$ has all possible paths of length $2$ from $v$ to $w$ as its neighbors (see~\cref{fig:related:wl-neighbors:2}).
Also note that, even though $k$-\acs{wl} refines $k$-tuple colors, lower-dimensional structures still get their own colors since a tuple does not have to consist of distinct vertices, i.e.\ in $k$-\acs{wl} the color of a single vertex $v \in \mathcal{V}_G$ is described by $\chi^{*}_{G, k}(s)$ for $s = (v, \dots, v) \in \mathcal{V}_G^k$.

Let us now look at how the tuple colors are initialized.
For this we use subgraph \textit{isomorphism types} which determine the initial color $\chi_{G,k}^{(0)}(v)$ of each $k$-tuple $v$.
For $k = 1$ the isomorphism type of a vertex $v$ directly corresponds to its label $l_G(v)$.
For $k > 1$ the isomorphism types more generally correspond to the classes of isomorphic ordered subgraphs of $G$ induced by the $k$-tuples $\mathcal{V}_G^k$.
Formally this means that
\begin{align}
	\chi_{G,k}^{(0)}(v_1, \dots, v_k) = \chi_{G,k}^{(0)}(w_1, \dots, w_k) \label{eq:related:wlk-init}
	\iff &\ \forall i: l_G(v_i) = l_G(w_i) \\
	\land &\ \forall i, j: v_i = v_j \leftrightarrow w_i = w_j \nonumber \\
	\land &\ \forall i, j: (v_i, v_j) \in \mathcal{E}_G \leftrightarrow (w_i, w_j) \in \mathcal{E}_G \text{.} \nonumber
\end{align}
If no explicit vertex labeling $l_G$ is given, a single constant label for all vertices is assumed.
Also note that there is a fundamental difference in how the adjacency information encoded in $\mathcal{E}_G$ is used in 1-\acs{wl} vs.\ $k$-\acs{wl}:
In 1-\acs{wl} a vertex coloring by itself cannot encode adjacency which is why this information is explicitly incorporated in each refinement step via $\Gamma_G$ (see~\cref{defn:related:wl1-refine}).
In $k$-\acs{wl} on the other hand each pair of vertices $(v, u) \in \mathcal{V}_G^2$ appears in at least one $k$-tuple (assuming $k \geq 2$) and therefore has at least one color which can implicitly encode the adjacency information.
Edges and non-edges are colored differently in the initial coloring as defined in \cref{eq:related:wlk-init};
thus no explicit adjacency information is needed in the $k$-\acs{wl} color refinement step in \cref{defn:related:wlk-refine}.

\subsubsection{Discriminative Power of \acs{wl}}
Now we will look at the types of graphs that can be distinguished by \acs{wl} in relation to the \acs{wl}-dimension $k$.
\begin{lem}\label{lem:related:wl-lower}
	$G \mathrel{\not\simeq_k} H \implies G \mathrel{\not\simeq_{k+1}} H$ ($(k+1)$-\acs{wl} is at least as powerful as $k$-\acs{wl}).
\end{lem}
\begin{hproof}
	All $k$-tuples can be mapped to $(k+1)$-tuples via $\varphi(v_1, \dots, v_k) \coloneqq (v_1, \dots,\allowbreak v_k, v_k)$.
	For each neighbor $(v[u/1], \dots,\allowbreak v[u/k])$ of $v$, there is a corresponding neighbor $(\varphi(v[u/1]), \dots \varphi(v[u/k]), \varphi(v[u/k]))$ of $\varphi(v)$.
	Using \cref{eq:related:wlk-init} (for $i = 0$) and \cref{defn:related:wlk-refine} (for $i > 0$) it follows that $\forall v, w \in \mathcal{V}_G^k: \chi_{G, k}^{(i)}(v) \neq \chi_{G, k}^{(i)}(w) \rightarrow \chi_{G, k+1}^{(i)}(\varphi(v)) \neq \chi_{G, k+1}^{(i)}(\varphi(w))$.
	The lemma then follows by \cref{defn:related:wl-distinguishable}.
\end{hproof}
\begin{prop}[see~\citet{Immerman1990}]\label{prop:related:wl-upper}
	For all $k \in \mathbb{N}$ there are non-isomorphic graphs $G \not\simeq H$ of size $\mathcal{O}(k)$ with $G \mathrel{\simeq_k} H$.
\end{prop}
\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{gfx/related-work/wl2-problem-solution.pdf}
	\caption{
		Two non-isomorphic graphs with $G \mathrel{\simeq_1} H$ and $G \mathrel{\not\simeq_2} H$.
	}\label{fig:related:wl2-problem-solution}
\end{figure}
\begin{cor}\label{cor:related:wl-monotonous}
	The discriminative power of $k$-\acs{wl} grows monotonously with $k$ and never converges, i.e.\ for all $k \in \mathbb{N}$ there is an $l \in \mathbb{N}$ s.t.\ the set of $k$-\ac{wl} distinguishable graphs is a proper subset of the $(k+l)$-\ac{wl} distinguishable graphs.
\end{cor}
\begin{hproof}
	Note that $k$-\acs{wl} trivially solves the graph isomorphism problem for all graphs of size $\leq k$ via the initial coloring (see~\cref{eq:related:wlk-init}).
	Using \cref{lem:related:wl-lower} and \cref{prop:related:wl-upper} the corollary directly follows.
\end{hproof}
\Cref{fig:related:wl2-problem-solution} illustrates \cref{cor:related:wl-monotonous} by showing how 2-WL is able to distinguish the two 1-WL indistinguishable graphs from \cref{fig:related:wl1-problem}.
Since the time complexity of $k$-\acs{wl} grows exponentially with $k$~\cite[corr.~1.9.7]{Immerman1990}, a trade-off between the runtime and the discriminative power has to be made.

@@

\paragraph{Graph Embeddings}
Skip-gram methods can not only be used for vertex embeddings but also to embed entire graphs.
One way to do this is via the \texttt{graph2vec}~\cite{Narayanan2017} method.
It is inspired by \texttt{doc2vec}~\cite{Le2014} which in turn is based on \texttt{word2vec}.
\texttt{graph2vec} gets a set of graphs $\mathbb{G} = \{ G_1, \dots G_N \}$ as input and outputs graph embeddings $f(G_1), \dots, f(G_N)$.
While in \texttt{word2vec} every word can be a context root as well as a context element, \texttt{graph2vec} uses the graphs $\mathbb{G}$ as context roots.
The context $C_k(G_i)$ of a graph $G_i$ is defined as
\begin{equation}
	C_k(G_i) \coloneqq \bigcup_{\mathclap{v_j \in \mathcal{V}_{G_i}}}\, {\left\{ {\chi_{G_i, 1}^{(l)}(v_j) \right\}}_{l = 0}^k % chktex 21
	\text{ with $k \in \mathbb{N}$ and $\chi_{G_i, 1}^{(l)}$ as in \cref{defn:related:wl1-refine}.}
\end{equation}
Intuitively this context can be understood as the set of 1-\acs{wl}-distinguishable subgraphs of $G_i$ with diameter $\leq 2k$.
Since \ac{wl} is used to identify distinct subgraphs, \texttt{graph2vec} can only be applied to graphs with discrete vertex labels.
Using the previous definitions, the context root embedding function has the signature $f: \mathbb{G} \to \mathbb{R}^d$ while the context element embedding function is of type $f_C: \bigcup_{G_i \in \mathbb{G}} C_k(G_i) \to \mathbb{R}^d$.
To find those embeddings the \texttt{word2vec} objective from \cref{eq:related:gcr:embed:word2vec} is reused.
Analogous to \texttt{word2vec}, \texttt{graph2vec} therefore embeds graphs that share subgraphs close to each other, whereas graphs that do not share substructures tend to be embedded further away from each other.
